Compositional Equivalences Based on Open pNets

### 2022-10-17 2016-Weak-FH-Bisim/WeakBisim.tex
similarity with symbolic bisimulations
en page 4
comments reu

Bisimulation over a symbolic and open model like open pNets or open automata is  different from the classical notion of bisimulation because it cannot rely on the equality over a finite set of action labels. Classical bisimulations require to exhibit, for each transition of one system, a transition of the other system that simulates it. Instead, bisimulation for open automata  relies on the simulation of each open transition of one automaton by \emph{a set of} open transitions of the other one, that should cover all the 
cases where the original transition can be triggered. \TODO{Eric: is this like in symbolic bisimulation ?}

@@

Two open automata are FH-bisimilar if there exists an FH-bisimulation that relates their initial states\footnote{I.e., the predicate relation associated to the initial states is \True.}. We call this relation \emph{FH-bisimilarity}.

@@

\begin{example}
The simulation of one transition by many others is one non-standard aspect of this definition. This is made necessary by the parameterised nature of our model. Consider the open transition:
\[ \openrule
         {
           \set{\beta}, \True,\subst{y\leftarrow x}}
          {s \OTarrow {\alpha(x)} s'}\]
Bisimulation should allow it to be matched by the two following ones (depending on the value of $x$):
\[ \openrule
         {
           \set{\beta}, x\geq 0,\subst{y\leftarrow x}}
          {s \OTarrow {\alpha(x)} s'}
\qquad
\openrule
         {
           \set{\beta}, x<0,\subst{y\leftarrow x}}
          {s \OTarrow {\alpha(x)} s'}
\]
This example illustrates the necessity of multiple transitions in the definition of bisimulation in a naive and minimalistic way.  It can easily be extended into a non-trivial example with more states and different usage of the variables.
\end{example}

>>>>

Bisimulation over a symbolic and open model like open pNets or open automata is  different from the classical notion of bisimulation because it cannot rely on the equality over a finite set of action labels. Classical bisimulations require to exhibit, for each transition of one system, a transition of the other system that simulates it. Instead, bisimulation for open automata  relies on the simulation of each open transition of one automaton by \emph{a set of} open transitions of the other one, that should cover all the 
cases where the original transition can be triggered. %\TODO{Eric: is this like in symbolic bisimulation ?}
\ERIC{This is similar to the early and late symbolic bisimulation equivalences for value-passing CCS \cite{HENNESSY:1995}, though we need more general definitions in our setting.}

@@

Two open automata are FH-bisimilar if there exists an FH-bisimulation that relates their initial states\footnote{In other words, the predicate relation associated to the initial states is \True.}. We call this relation \emph{FH-bisimilarity}.

@@

\begin{example}
The simulation of one transition by many others is one non-standard aspect of this definition. This is made necessary by the parameterised nature of our model. Consider the following open transition.
\[ \openrule
         {
           \set{\beta}, \True,\subst{y\leftarrow x}}
          {s_1 \OTarrow {\alpha(x)} s_1'}\]
Bisimulation should allow it to be matched by the two following ones (depending on the value of $x$), to prove the bisimulation $\mathcal{R}=\{(s_1,s_2,\True),(s_1',s_2',\True)\}$
\[ \openrule
         {
           \set{\beta}, x\geq 0,\subst{y\leftarrow x}}
          {s_2 \OTarrow {\alpha(x)} s_2'}
\qquad
\openrule
         {
           \set{\beta}, x<0,\subst{y\leftarrow x}}
          {s_2 \OTarrow {\alpha(x)} s_2'}
\]
This example illustrates the necessity of multiple transitions in the definition of bisimulation in a naive and minimalistic way.  It can easily be extended into a non-trivial example with more states and different usage of the variables.
\end{example}
